---
title: "606 Chapter 3 Lab"
author: "Munkhnaran Gankhuyag"
date: "September 17, 2017"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load-data, eval=TRUE}
load("more/bdims.RData")
```
```{r male-female, eval=TRUE}
mdims <- subset(bdims, sex == 1)
fdims <- subset(bdims, sex == 0)

head(bdims)
```
## The normal distribution

1.  Make a histogram of men's heights and a histogram of women's heights.  How 
    would you compare the various aspects of the two distributions?

Based on the histograms, most frequent height for men and women is around 178 and 158. 
```{r}
hist(mdims$hgt, breaks = 50,  main = "Histogram of Men Heights in cm")
hist(fdims$hgt, breaks = 50, main = "Histogram of Women Heights in cm")
```

```{r female-hgt-mean-sd, eval=TRUE}
fhgtmean <- mean(fdims$hgt)
fhgtsd   <- sd(fdims$hgt)
fwgtmean <- mean(fdims$wgt)
fwgtsd   <- sd(fdims$wgt)

fhgtmean
fhgtsd
fwgtmean
fwgtsd
```
```{r hist-height, eval=TRUE}
hist(fdims$hgt, probability = TRUE, ylim = c(0, 0.06))
x <- 140:190
y <- dnorm(x = x, mean = fhgtmean, sd = fhgtsd)
lines(x = x, y = y, col = "blue")
```
2.  Based on the this plot, does it appear that the data follow a nearly normal 
    distribution?
    
    Yes, based on the plot it appears to follow a normal distribution.
    
## Evaluating the normal distribution
```{r qq, eval=TRUE}
qqnorm(fdims$hgt)
qqline(fdims$hgt)
```

```{r sim-norm, eval=TRUE}
sim_norm <- rnorm(n = length(fdims$hgt), mean = fhgtmean, sd = fhgtsd)
```    
    
3.  Make a normal probability plot of `sim_norm`.  Do all of the points fall on 
    the line?  How does this plot compare to the probability plot for the real 
    data?
    
    
```{r}
qqnorm(sim_norm)
qqline(sim_norm)

```
the simulated model is similar to the probability plot of the real data where most of the data is on the line except at the tails.
```{r qqnormsim, eval=TRUE}
qqnormsim(fdims$hgt)
```

4.  Does the normal probability plot for `fdims$hgt` look similar to the plots 
    created for the simulated data?  That is, do plots provide evidence that the
    female heights are nearly normal?
    
    Yes, the normal probablity plot for female heights look similar the plots created for the simulated data. Its follows the line except for at the tails of the line.
    
```{r, eval=TRUE}
F_Weight <- qqnormsim(fdims$wgt)
F_Weight
```

5.  Using the same technique, determine whether or not female weights appear to 
    come from a normal distribution.
    
Female weight also appear to come from a normal distribution except at the tails but, the simulations do not have the outliers that are in the real data .

## Normal probabilities

```{r pnorm, eval=TRUE}
1 - pnorm(q = 182, mean = fhgtmean, sd = fhgtsd)
sum(fdims$hgt > 182) / length(fdims$hgt)
```
6.  Write out two probability questions that you would like to answer; one 
    regarding female heights and one regarding female weights.  Calculate the 
    those probabilities using both the theoretical normal distribution as well 
    as the empirical distribution (four probabilities in all).  Which variable,
    height or weight, had a closer agreement between the two methods?

What is the probablity that randomly chosen female is taller than (5'11) 180 cm?

```{r}
1 - pnorm(q = 180, mean = fhgtmean, sd =fhgtsd)
sum(fdims$hgt > 180) / length(fdims$hgt)
```

What is the probablity that randomly chosen female is under 175 lbs?
```{r}


pnorm(q = 79, mean = fwgtmean, sd =fwgtsd)
sum(fdims$wgt < 79) / length(fdims$wgt)
```

Between the probability models, the weight had a closer agreement between the two methods. The theoretical normal distribution predicts that there are 97.2% chance that a random chosen female is under 175 lbs while the empirical distribution says that there are is a 95.4% chance.

## On Your Own


-   Now let's consider some of the other variables in the body dimensions data 
    set.  Using the figures at the end of the exercises, match the histogram to 
    its normal probability plot.  All of the variables have been standardized 
    (first subtract the mean, then divide by the standard deviation), so the 
    units won't be of any help.  If you are uncertain based on these figures, 
    generate the plots in R to check.


```{r}
fdims[1:20 ,]
hist(fdims$bii.di, breaks = 20)
```

    **a.** The histogram for female biiliac (pelvic) diameter (`bii.di`) belongs
    to normal probability plot letter D .

    **b.** The histogram for female elbow diameter (`elb.di`) belongs to normal 
    probability plot letter C.

    **c.** The histogram for general age (`age`) belongs to normal probability 
    plot letter B.

    **d.** The histogram for female chest depth (`che.de`) belongs to normal 
    probability plot letter A.

-   Note that normal probability plots C and D have a slight stepwise pattern.  
    Why do you think this is the case?
    
The are measurements of diameters of the pelvic and elbow, and there they're all around the same size.

-   As you can see, normal probability plots can be used both to assess 
    normality and visualize skewness.  Make a normal probability plot for female 
    knee diameter (`kne.di`).  Based on this normal probability plot, is this 
    variable left skewed, symmetric, or right skewed?  Use a histogram to confirm 
    your findings.

The normal probability plot for female knee diameter looks like it is skewed right.
    
```{r, eval=TRUE}
qqnorm(fdims$kne.di)
qqline(fdims$kne.di)
```
    
```{r hist-knee diameter, eval=TRUE}
range(fdims$kne.di)
fknee.dimean <- mean(fdims$kne.di)
fknee.disd <-sd(fdims$kne.di)

hist(fdims$kne.di, probability = TRUE,)
x <- 1:100
y <- dnorm(x = x, mean = fknee.dimean, sd = fknee.disd)
lines(x = x, y = y, col = "green")
```
  
  Based on the historgram, it looks like the data is skewed right.  

    
